Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-03-15 introduction.tex

The numerical simulation of astronomical phenomena now plays a key role, along with observations and analytic theory, in pushing forward our understanding of the cosmos.
%\citep[e.g.,][]{DecadalSurvey01}.  
But we believe that with this role comes responsibility.  Those developing simulation tools must fulfill two key obligations: the first is to make those tools available to the community as a whole, much in the way that astronomical data is usually made publicly available.  The second is document, test and refineme those methods so that they can be critically evaluated and expanded by others.  This paper represents our attempt to meet our obligations with respect to the \enzo\ code.

@@

There are many diverse phenomena in astrophysics and cosmology which require a high spatial and temporal dynamical range. There are a number of numerical techniques which have been used to
model such problems.  The most common is a gridless, fully Lagrangian method call Smoothed Particle Hydrodynamcs \citep[SPH;][]{Lucy77, SPH} which has been very successful, particularly for those situations dominated by gravity.  Its development, however, is still at a relatively early stage when
compared to the effort put into Eulerian, grid-based hydrodynamic schemes \citep{Woodward84}.

Therefore, we would like to make use of this expertise; unfortunately, none of these schemes provide an easy way to adaptively and dynamically increase the resolution in small volumes, as required by, for example, gravitational instability.  We have instead, adapted an idea first suggested by \citet{Berger89} in the Computational Fluid Dynamics (CFD) community, which has become known as Structured Adaptive Mesh Refinement (SAMR).  The principle is to adaptively add (and modify) additional, finer meshes, or patches, over regions which require higher resolution.  Since each patch has a simple geometry (here, cartesian) the hyperbolic fluid-dynamical equations can be solved on that patch with a well-developed `off-the-shelf' numerical method.  In addition, it will be relatively straightforward to add other advanced physics (e.g. MHD, radiation transport beyond flux limited diffusion, etc.). This paper is devoted to describing how we have implemented this idea for astrophysics, including the addition of self-gravity and collisionless fluids.

There have been a few other numerical methods described in the astronomical literature which contain elements of this idea, or have a similar aim.  For example, an N-body solver \cite{Villumsen89} used
non-adaptive meshing to increase the resolution in pre-selected.  This static approach, applied to hydrodynamics, has been used extensively \citep[e.g.,]{Ruffert94, Anninos94}.  Adding adaptivity is a more recent enhancement, and there are now a number of N-body codes which possess this feature, both with and without  hydrodynamics  \cite{Couchman91, Jessop94, Suisalu95, Splinter96, Gelato97, ART97, Truelove98, flash_method, MLAPM01,  Yahagi01, RAMSES, Quilis04, Ziegler05, Zhang06}. It is also possible to deform the grid to obtain high resolution \citep[e.g.,][]{Gnedin95, Xu97, Pen98}.

The \enzo\ code has been widely used over the lasts decade in a wide variety of problems in topics ranging from star formation to turbulence to large-scale structure.  More than 50 peer-reviewed papers have been published based on results obtained with the code.

>>>>

There are many diverse phenomena in astrophysics and cosmology which require a high spatial and temporal dynamical range. There are a number of numerical techniques which have been used to
model such problems.  The most common is a gridless, fully Lagrangian method call Smoothed Particle Hydrodynamcs \citep[SPH;][]{Lucy77, SPH} which has been very successful, particularly for those situations dominated by gravity.  Its development, however, is still at a relatively early stage when
compared to the effort put into Eulerian, grid-based hydrodynamic schemes \citep[e.g.,][]{Woodward84}.

Therefore, we would like to make use of this expertise; unfortunately, none of these schemes provide an easy way to adaptively and dynamically increase the resolution in small volumes, as required by, for example, gravitational instability.  We have instead, adapted an idea first suggested by \citet{Berger89} in the Computational Fluid Dynamics (CFD) community, which has become known as Structured Adaptive Mesh Refinement (SAMR).  The principle is to adaptively add (and modify) additional, finer meshes, or patches, over regions which require higher resolution.  Since each patch has a simple geometry (here, cartesian) the hyperbolic fluid-dynamical equations can be solved on that patch with a well-developed `off-the-shelf' numerical method.  In addition, it is relatively straightforward to add other advanced physics (e.g. MHD, radiation transport, etc.). This paper is devoted to describing how we have implemented this idea for astrophysics, including the addition of self-gravity, collisionless fluids, MHD, radiation transport and a range of other physical effects.

There have been many other numerical methods described in the astronomical literature which contain elements of this idea, or have a similar aim.  For example, an N-body solver \cite{Villumsen89} used non-adaptive meshing to increase the resolution in pre-selected regions.  This static approach, applied to hydrodynamics, has been used extensively \citep[e.g.,]{Ruffert94, Anninos94}.  Adding adaptivity is a more recent enhancement, and there are now a number of N-body codes which possess this feature, both with and without  hydrodynamics  \cite{Couchman91, Jessop94, Suisalu95, Splinter96, Gelato97, ART97, Truelove98, flash_method, MLAPM01,  Yahagi01, RAMSES, Quilis04, Ziegler05, Zhang06}.  Of these, perhaps the most comparable are FLASH \citep{flash_method}, which uses grid blocks of fixed size, and RAMSES \citep{RAMSES} and ART \citep{ART97}, both of which refine individual cells.  It is also possible to deform the grid to obtain high resolution \citep[e.g.,][]{Gnedin95, Xu97, Pen98}, and more recently

@@

The \enzo\ code has been widely used over the lasts decade in a wide variety of problems in topics ranging from star formation to turbulence to large-scale structure.  More than 100 peer-reviewed papers have been published based on results obtained with the code.

Finally, we note that the numerical simulation of astronomical phenomena now plays a key role, along with observations and analytic theory, in pushing forward our understanding of the cosmos.
%\citep[e.g.,][]{DecadalSurvey01}.  
But along with this role comes responsibility.  We believe that those developing simulation tools must fulfill two key obligations: the first is to make those tools available to the community as a whole, much in the way that astronomical data is now regularly made publicly available.  The second is document, test and refine those methods so that they can be critically evaluated and expanded by others.  This paper represents one part of our attempt to meet our obligations with respect to the \enzo\ code.
